---
title: "1_tables"
output: html_document
---

Purpose: Generate supplemental and regression tables

Date created: February 23, 2026

**Note: this RMD was not fully formatted to be knitted, but rather for retrieving and saving relevant tables for this manuscript.**

```{r setup, include=FALSE}

library(tidycensus)
library(qwraps2)
library(tigris)
library(dplyr)
library(stringr)
library(sf)
library(here)
library(tidyverse)

# pull in data
exposure_data_full <- read.csv(here("data", "processed_data", "zip3_exposure_dataset.csv"))

env_data <- read.csv(here("data", "processed_data", "env_data.csv"))

demodata <- read.csv(here("data", "processed_data", "demodata.csv"))

# subsetting exposure_data to just the 2024 year (refers only to population), so that we have a dataframe with exposure for each zip3
exposure_data <- subset(exposure_data_full, subset = exposure_data_full$year == "2024")

# join exposure data and demographic data
exposure_data$zip3 <- as.character(exposure_data$zip3)
demodata$zip3 <- as.character(demodata$zip3)

table_data <- full_join(demodata, exposure_data, by = "zip3")

# creating new column labeling exposure to flooding for table columns
table_data <- table_data |> 
  mutate(exposure_desc = if_else(inundation_exposure == TRUE, 
                                 "Exposed to Flooding", 
                                 "Not Exposed to Flooding")
         )

# analytic datasets
dat <- read.csv(here("data", "processed_data", "analytic_dataset.csv"))
dat_all_imputed <- read.csv(here("data", "processed_data", "analytic_imputed_datasets.csv"))

```

# Table S1: demographic comparison between exposure groups

```{r, results='asis'}
options(qwraps2_markup = "markdown")

# cleans table data
table_data_c <- table_data %>%
  st_drop_geometry() %>%
  mutate(inundation_exposure = as.factor(inundation_exposure)) %>%
  as.data.frame()

# template for table
demographic_stats <- 
  list("Race and Ethnicity" = 
         list("Total Population" = ~ format(round(sum(estimate_total_population, na.rm = TRUE), -2),
                                            big.mark = ","), 
              "non-Hispanic Black" = ~paste0(format(round(sum(estimate_black, na.rm = TRUE), -2),
                                                    big.mark = ","), 
                                             " (", round(sum(estimate_black, na.rm = TRUE) / 
                                                           sum(estimate_total_population, na.rm = TRUE)*100, 1), "%)"), 
              "non-Hispanic White" = ~paste0(format(round(sum(estimate_white, na.rm = TRUE), -2),
                                                    big.mark = ","), 
                                             " (", round(sum(estimate_white, na.rm = TRUE) / 
                                                           sum(estimate_total_population, na.rm = TRUE)*100, 1), "%)"), 
              "non-Hispanic Asian" = ~paste0(format(round(sum(estimate_asian, na.rm = TRUE), -2),
                                                          big.mark = ","), 
                                             " (", round(sum(estimate_asian, na.rm = TRUE) / 
                                                           sum(estimate_total_population, na.rm = TRUE)*100, 1), "%)"),
              "non-Hispanic American Indian or Alaskan Native" = ~paste0(format(round(sum(estimate_AI_AN, na.rm = TRUE), -2),
                                                                                big.mark = ","), 
                                             " (", round(sum(estimate_AI_AN, na.rm = TRUE) / 
                                                           sum(estimate_total_population, na.rm = TRUE)*100, 1), "%)"),
              "non-Hispanic Two or More Races" = ~paste0(format(round(sum(estimate_two_more, na.rm = TRUE), -2),
                                                                big.mark = ","), 
                                             " (", round(sum(estimate_two_more, na.rm = TRUE) / 
                                                           sum(estimate_total_population, na.rm = TRUE)*100, 1), "%)"),
              "Hispanic" = ~paste0(format(round(sum(estimate_hispanic, na.rm = TRUE), -2),
                                          big.mark = ","), 
                                             " (", round(sum(estimate_hispanic, na.rm = TRUE) / 
                                                           sum(estimate_total_population, na.rm = TRUE)*100, 1), "%)")), 
  "Other Demographic Factors" = 
         list("Median Household Income [USD]" = ~mean_sd(avg_med_income, digits = 0), 
              "Number of Rural Households" = ~paste0(format(round(sum(rural_hhs, na.rm = TRUE), -2),
                                                            big.mark = ","), 
                                                     " (", round(sum(rural_hhs, na.rm = TRUE) / 
                                                                   sum(total_hhs, na.rm = TRUE)*100, 1), "%)"),
              "Health Insurance" = ~paste0(format(round(sum(estimate_uninsured, na.rm = TRUE), -2),
                                                  big.mark = ","), 
                                              " (", round(sum(estimate_uninsured, na.rm = TRUE) / 
                                                            sum(estimate_total_ins, na.rm =TRUE) *100, 1), "%)"), 
              "Area [sqmi]" = ~format(round(sum(as.numeric(area_sqmi), na.rm = TRUE), -1), big.mark = ","), 
              "Population Density [People per sqmi]" = ~round(sum(estimate_total_population, na.rm = TRUE) / 
                                                                sum(area_sqmi, na.rm = TRUE), 0)))


summary_table(table_data_c, demographic_stats)

# creates table using the cleaned dataframe, specifies the columns to be exposed/unexposed
summary_table(table_data_c, demographic_stats, by = "exposure_desc")

 
```

# Table 1: regression results from primary models

```{r}
# assemble table 1
results_files <- list.files(here("tables"))

main_results_files <- results_files[
  grepl(paste(c("main_model","imputation_results"), collapse = "|"), 
        results_files)]

main_results_files <- paste("../../tables/", main_results_files, sep = "")

main_results <- read.csv(main_results_files[2])
pooled_results <- read.csv(main_results_files[1])

main_results <- main_results %>%
  dplyr::select(model_group, model_type, plot_estimate, nobs) %>%
  mutate(total_events = sum(dat$n_events))

total_imputed_events <- dat_all_imputed %>% 
  group_by(imputation) %>% 
  summarise(n = sum(completed_events)) %>% 
  ungroup()

pooled_results <- pooled_results %>%
  dplyr::select(model_group, model_type, plot_estimate) %>%
  mutate(nobs = unique(table(dat_all_imputed$imputation)),
         total_events = round(median(total_imputed_events$n), 0))

table_one <- rbind(main_results, pooled_results)

if (FALSE) {
  write.csv(table_one, here("regression_results", "tables", "table1.csv"))
}

```

# Analyses focused on private wells


```{r}
# assemble table 1
results_files <- list.files(here("tables"))

main_results_files <- results_files[
  grepl(paste(c("main_model","imputation_results"), collapse = "|"), results_files)]

main_results_files <- paste("../../tables/", main_results_files, sep = "")

main_results <- read.csv(main_results_files[2])
pooled_results <- read.csv(main_results_files[1])

main_results <- main_results %>%
  dplyr::select(model_group, model_type, plot_estimate, nobs) %>%
  mutate(total_events = sum(dat$n_events))

total_imputed_events <- dat_all_imputed %>% 
  group_by(imputation) %>% 
  summarise(n = sum(completed_events)) %>% 
  ungroup()

pooled_results <- pooled_results %>%
  dplyr::select(model_group, model_type, plot_estimate) %>%
  mutate(nobs = unique(table(dat_all_imputed$imputation)),
         total_events = round(median(total_imputed_events$n), 0))

table_one <- rbind(main_results, pooled_results)

if (FALSE) {
  write.csv(table_one, here("tables", "table1.csv"))
}

```


- results for private well analyses
- sensitivity analyses (figure or table)

# Hurricane Florence results

```{r}
florence_results <- read.csv(paste("../../tables/", results_files[1], sep = ""))

florence_results_table <- florence_results %>%
  dplyr::select(model_group, model_type, plot_estimate, nobs)

if (FALSE) {
  write.csv(florence_results_table, "../../tables/tableS2.csv")
}

```

# Graphical abstract
```{r}
# TODO create an example CITS figure?

CITS_dat <- tibble(
  t = c(1:50, 1:50),
  group = c(rep("Exposed", 50), rep("Unexposed", 50)),
  event = ifelse(t < 25, 0, )
  outcome = ifelse(
    group == "Exposed", rpois(n = 50, 2), rpois(n = 50, 2 + 0.5)
    
  )
)

```

```{r}

set.seed(42)

# 1. Define study parameters
n_timepoints <- 40
intervene_at <- 21  # Intervention occurs at time point 21

# 2. Build full factorial grid (Time x Group)
df <- expand.grid(
  time = 1:n_timepoints,
  group = c("Control", "Treatment")
) %>%
  mutate(
    group_num  = ifelse(group == "Treatment", 1, 0),
    post       = ifelse(time >= intervene_at, 1, 0),
    time_after = ifelse(post == 1, time - intervene_at + 1, 0)
  )

# 3. Define log-linear coefficients (log scale for count data)
beta_0 <- 2.50   # Baseline log count (Control at t = 0)
beta_1 <- 0.01   # Pre-intervention slope (Control)
beta_2 <- 0.20   # Baseline group difference (Treatment vs Control)
beta_3 <- 0.005  # Pre-intervention slope difference
beta_4 <- -0.05  # Control level change post-intervention (external shock)
beta_5 <- -0.002 # Control slope change post-intervention
beta_6 <- 0.40   # Immediate intervention level change (Treatment effect)
beta_7 <- 0.035  # Sustained intervention slope change (Treatment effect)

# 4. Compute expected mean (mu) and generate Poisson count outcome
df <- df %>%
  mutate(
    log_mu = beta_0 + 
             beta_1 * time + 
             beta_2 * group_num + 
             beta_3 * (time * group_num) + 
             beta_4 * post + 
             beta_5 * time_after + 
             beta_6 * (post * group_num) + 
             beta_7 * (time_after * group_num),
    
    mu = exp(log_mu),
    
    # Generate outcome Y ~ Poisson(mu)
    count = rpois(n(), lambda = mu)
  )

# 5. Fit Poisson GLM to estimate parameters
fit <- glm(
  count ~ time * group_num + post * group_num + time_after * group_num,
  family = poisson(link = "log"),
  data = df
)

# Add predicted values to dataframe for smooth plotting
df$pred <- predict(fit, type = "response")

# 6. Plot Controlled Interrupted Time Series
ggplot(df, aes(x = time, y = count, color = group, group = group)) +
  geom_point(alpha = 0.6, size = 2) +
  # Model fit trends pre and post-intervention
  geom_line(aes(y = pred), size = 1) +
  # Intervention vertical indicator line
  geom_vline(xintercept = intervene_at - 0.5, linetype = "dashed", color = "black") +
  annotate("text", x = intervene_at - 1, y = max(df$count), 
           label = "Intervention", hjust = 1, fontface = "italic") +
  scale_color_manual(values = c("Control" = "#2b5c8f", "Treatment" = "#d95f02")) +
  labs(
    title = "Controlled Interrupted Time Series (CITS)",
    subtitle = "Simulated Poisson Count Outcome",
    x = "Time Period",
    y = "Count Outcome",
    color = "Group"
  ) +
  theme_minimal(base_size = 12) +
  theme(legend.position = "top")

```


