---
title: "1_tables"
output: html_document
---

Purpose: Generate supplemental and regression tables

Date created: February 23, 2026

**Note: this RMD was not fully formatted to be knitted, but was only for retrieving and saving relevant tables for this manuscript.**

```{r setup, include=FALSE}
library(tidycensus)
library(qwraps2)
library(tigris)
library(dplyr)
library(stringr)
library(sf)

# pull in data
exposure_data_full <- read.csv("../../data/processed_data/zip3_exposure_dataset.csv")

env_data <- read.csv("../../data/processed_data/env_data.csv")

demodata <- read.csv("../../data/processed_data/demodata.csv")

# subsetting exposure_data to just the 2024 year (refers only to population), so that we have a dataframe with exposure for each zip3
exposure_data <- subset(exposure_data_full, subset = exposure_data_full$year == "2024")

# join exposure data and demographic data
exposure_data$zip3 <- as.character(exposure_data$zip3)
demodata$zip3 <- as.character(demodata$zip3)

table_data <- full_join(demodata, exposure_data, by = "zip3")

# creating new column labeling exposure to flooding to clean up table columns
table_data <- table_data |> mutate(exposure_desc = if_else(inundation_exposure == TRUE, "Exposed to Flooding", "Not Exposed to Flooding"))

# analytic datasets
dat <- read_csv("../../data/processed_data/analytic_dataset.csv")
dat_all_imputed <- read_csv("../../data/processed_data/analytic_imputed_datasets.csv")

```

# Table S1: demographic comparison between exposure groups

```{r, results='asis'}
options(qwraps2_markup = "markdown")

# cleans table data
table_data_c <- table_data %>%
  st_drop_geometry() %>%
  mutate(inundation_exposure = as.factor(inundation_exposure)) %>%
  as.data.frame()

# template for table
demographic_stats <- 
  list("Race and Ethnicity" = 
         list("Total Population" = ~ format(round(sum(estimate_total_population, na.rm = TRUE), -2),
                                            big.mark = ","), 
              "non-Hispanic Black" = ~paste0(format(round(sum(estimate_black, na.rm = TRUE), -2),
                                                    big.mark = ","), 
                                             " (", round(sum(estimate_black, na.rm = TRUE) / 
                                                           sum(estimate_total_population, na.rm = TRUE)*100, 1), "%)"), 
              "non-Hispanic White" = ~paste0(format(round(sum(estimate_white, na.rm = TRUE), -2),
                                                    big.mark = ","), 
                                             " (", round(sum(estimate_white, na.rm = TRUE) / 
                                                           sum(estimate_total_population, na.rm = TRUE)*100, 1), "%)"), 
              "non-Hispanic Asian" = ~paste0(format(round(sum(estimate_asian, na.rm = TRUE), -2),
                                                          big.mark = ","), 
                                             " (", round(sum(estimate_asian, na.rm = TRUE) / 
                                                           sum(estimate_total_population, na.rm = TRUE)*100, 1), "%)"),
              "non-Hispanic American Indian or Alaskan Native" = ~paste0(format(round(sum(estimate_AI_AN, na.rm = TRUE), -2),
                                                                                big.mark = ","), 
                                             " (", round(sum(estimate_AI_AN, na.rm = TRUE) / 
                                                           sum(estimate_total_population, na.rm = TRUE)*100, 1), "%)"),
              "non-Hispanic Two or More Races" = ~paste0(format(round(sum(estimate_two_more, na.rm = TRUE), -2),
                                                                big.mark = ","), 
                                             " (", round(sum(estimate_two_more, na.rm = TRUE) / 
                                                           sum(estimate_total_population, na.rm = TRUE)*100, 1), "%)"),
              "Hispanic" = ~paste0(format(round(sum(estimate_hispanic, na.rm = TRUE), -2),
                                          big.mark = ","), 
                                             " (", round(sum(estimate_hispanic, na.rm = TRUE) / 
                                                           sum(estimate_total_population, na.rm = TRUE)*100, 1), "%)")), 
  "Other Demographic Factors" = 
         list("Median Household Income [USD]" = ~mean_sd(avg_med_income, digits = 0), 
              "Number of Rural Households" = ~paste0(format(round(sum(rural_hhs, na.rm = TRUE), -2),
                                                            big.mark = ","), 
                                                     " (", round(sum(rural_hhs, na.rm = TRUE) / 
                                                                   sum(total_hhs, na.rm = TRUE)*100, 1), "%)"),
              "Health Insurance" = ~paste0(format(round(sum(estimate_uninsured, na.rm = TRUE), -2),
                                                  big.mark = ","), 
                                              " (", round(sum(estimate_uninsured, na.rm = TRUE) / 
                                                            sum(estimate_total_ins, na.rm =TRUE) *100, 1), "%)"), 
              "Area [sqmi]" = ~format(round(sum(as.numeric(area_sqmi), na.rm = TRUE), -1), big.mark = ","), 
              "Population Density [People per sqmi]" = ~round(sum(estimate_total_population, na.rm = TRUE) / 
                                                                sum(area_sqmi, na.rm = TRUE), 0)))


summary_table(table_data_c, demographic_stats)

# creates table using the cleaned dataframe, specifies the columns to be exposed/unexposed
summary_table(table_data_c, demographic_stats, by = "exposure_desc")

 
```


```{r}
# assemble table 2
results_files <- list.files("../../tables")

main_results_files <- results_files[
  grepl(paste(c("main_model","imputation_results"), collapse = "|"), results_files)]

main_results_files <- paste("../../tables/", main_results_files, sep = "")

main_results <- read.csv(main_results_files[2])
pooled_results <- read.csv(main_results_files[1])

main_results <- main_results %>%
  dplyr::select(model_group, model_type, plot_estimate, nobs) %>%
  mutate(total_events = sum(dat$n_events))

total_imputed_events <- dat_all_imputed %>% 
  group_by(imputation) %>% 
  summarise(n = sum(completed_events)) %>% 
  ungroup()

pooled_results <- pooled_results %>%
  dplyr::select(model_group, model_type, plot_estimate) %>%
  mutate(nobs = unique(table(dat_all_imputed$imputation)),
         total_events = round(median(total_imputed_events$n), 0))

table_one <- rbind(main_results, pooled_results)

if (FALSE) {
  write.csv(table_one, "../../tables/table1.csv")
}

```

# TODO: insert regression tables
- results for private well analyses
- senstivity analyses (figure or table)
- 
