---
title: "1_tables"
output: html_document
editor_options: 
  chunk_output_type: console
---

Purpose: Generate supplemental and regression tables

Date created: February 23, 2026

**Note: this RMD was not fully formatted to be knitted, but rather for retrieving and saving relevant tables for this manuscript.**

```{r setup, include=FALSE}

library(tidycensus)
library(qwraps2)
library(tigris)
library(dplyr)
library(stringr)
library(sf)
library(here)
library(tidyverse)

# pull in data
exposure_data_full <- read.csv(here("data", "processed_data", "zip3_exposure_dataset.csv"))

env_data <- read.csv(here("data", "processed_data", "env_data.csv"))

demodata <- read.csv(here("data", "processed_data", "demodata.csv"))

# subsetting exposure_data to just the 2024 year (refers only to population), so that we have a dataframe with exposure for each zip3
exposure_data <- subset(exposure_data_full, subset = exposure_data_full$year == "2024")

# join exposure data and demographic data
exposure_data$zip3 <- as.character(exposure_data$zip3)
demodata$zip3 <- as.character(demodata$zip3)

table_data <- full_join(demodata, exposure_data, by = "zip3")

# creating new column labeling exposure to flooding for table columns
table_data <- table_data |> 
  mutate(exposure_desc = if_else(inundation_exposure == TRUE, 
                                 "Exposed to Flooding", 
                                 "Not Exposed to Flooding")
         )

# analytic datasets
dat <- read.csv(here("data", "processed_data", "analytic_dataset.csv"))
dat_all_imputed <- read.csv(here("data", "processed_data", "analytic_imputed_datasets.csv"))

```

# Table S1: demographic comparison between exposure groups

```{r, results='asis'}
options(qwraps2_markup = "markdown")

# cleans table data
table_data_c <- table_data %>%
  st_drop_geometry() %>%
  mutate(inundation_exposure = as.factor(inundation_exposure)) %>%
  as.data.frame()

# template for table
demographic_stats <- 
  list("Race and Ethnicity" = 
         list("Total Population" = ~ format(round(sum(estimate_total_population, na.rm = TRUE), -2),
                                            big.mark = ","), 
              "non-Hispanic Black" = ~paste0(format(round(sum(estimate_black, na.rm = TRUE), -2),
                                                    big.mark = ","), 
                                             " (", round(sum(estimate_black, na.rm = TRUE) / 
                                                           sum(estimate_total_population, na.rm = TRUE)*100, 1), "%)"), 
              "non-Hispanic White" = ~paste0(format(round(sum(estimate_white, na.rm = TRUE), -2),
                                                    big.mark = ","), 
                                             " (", round(sum(estimate_white, na.rm = TRUE) / 
                                                           sum(estimate_total_population, na.rm = TRUE)*100, 1), "%)"), 
              "non-Hispanic Asian" = ~paste0(format(round(sum(estimate_asian, na.rm = TRUE), -2),
                                                          big.mark = ","), 
                                             " (", round(sum(estimate_asian, na.rm = TRUE) / 
                                                           sum(estimate_total_population, na.rm = TRUE)*100, 1), "%)"),
              "non-Hispanic American Indian or Alaskan Native" = ~paste0(format(round(sum(estimate_AI_AN, na.rm = TRUE), -2),
                                                                                big.mark = ","), 
                                             " (", round(sum(estimate_AI_AN, na.rm = TRUE) / 
                                                           sum(estimate_total_population, na.rm = TRUE)*100, 1), "%)"),
              "non-Hispanic Two or More Races" = ~paste0(format(round(sum(estimate_two_more, na.rm = TRUE), -2),
                                                                big.mark = ","), 
                                             " (", round(sum(estimate_two_more, na.rm = TRUE) / 
                                                           sum(estimate_total_population, na.rm = TRUE)*100, 1), "%)"),
              "Hispanic" = ~paste0(format(round(sum(estimate_hispanic, na.rm = TRUE), -2),
                                          big.mark = ","), 
                                             " (", round(sum(estimate_hispanic, na.rm = TRUE) / 
                                                           sum(estimate_total_population, na.rm = TRUE)*100, 1), "%)")), 
  "Other Demographic Factors" = 
         list("Median Household Income [USD]" = ~mean_sd(avg_med_income, digits = 0), 
              "Number of Rural Households" = ~paste0(format(round(sum(rural_hhs, na.rm = TRUE), -2),
                                                            big.mark = ","), 
                                                     " (", round(sum(rural_hhs, na.rm = TRUE) / 
                                                                   sum(total_hhs, na.rm = TRUE)*100, 1), "%)"),
              "Health Insurance" = ~paste0(format(round(sum(estimate_uninsured, na.rm = TRUE), -2),
                                                  big.mark = ","), 
                                              " (", round(sum(estimate_uninsured, na.rm = TRUE) / 
                                                            sum(estimate_total_ins, na.rm =TRUE) *100, 1), "%)"), 
              "Area [sqmi]" = ~format(round(sum(as.numeric(area_sqmi), na.rm = TRUE), -1), big.mark = ","), 
              "Population Density [People per sqmi]" = ~round(sum(estimate_total_population, na.rm = TRUE) / 
                                                                sum(area_sqmi, na.rm = TRUE), 0)))


summary_table(table_data_c, demographic_stats)

# creates table using the cleaned dataframe, specifies the columns to be exposed/unexposed
summary_table(table_data_c, demographic_stats, by = "exposure_desc")

```

# Table 1: regression results from primary models

```{r, echo=FALSE}
# assemble table 1
results_files <- list.files(here("regression_results"))

main_results_files <- results_files[
  grepl(paste(c("main_model","imputation_results"), collapse = "|"), 
        results_files)]

main_results <- read.csv(here("regression_results", main_results_files[2]))
pooled_results <- read.csv(here("regression_results", main_results_files[1]))

main_results <- main_results %>%
  dplyr::select(model_group, model_type, plot_estimate, nobs) %>%
  mutate(total_events = sum(dat$n_events))

total_imputed_events <- dat_all_imputed %>% 
  group_by(imputation) %>% 
  summarise(n = sum(completed_events)) %>% 
  ungroup()

pooled_results <- pooled_results %>%
  dplyr::select(model_group, model_type, plot_estimate) %>%
  mutate(nobs = unique(table(dat_all_imputed$imputation)),
         total_events = round(median(total_imputed_events$n), 0))

table_one <- rbind(main_results, pooled_results)

if (FALSE) {
  write.csv(table_one, here("regression_results", "tables", "table1.csv"))
}

```

# Sensitivity analyses for Hurricane Helene

```{r, echo=FALSE}

sensitivity_results_files <- results_files[
  grepl(paste(c("sensitivity_model","non-controlled_ITS_model"), collapse = "|"), 
        results_files)]

sensitivity_results <- read.csv(here("regression_results", sensitivity_results_files[2]))
ITS_results <- read.csv(here("regression_results", sensitivity_results_files[1]))

sensitivity_results <- sensitivity_results %>%
  dplyr::select(model_group, model_type, plot_estimate, nobs)

ITS_results <- ITS_results %>%
  dplyr::select(model_group, model_type, plot_estimate, nobs)

if (FALSE) {
  write.csv(sensitivity_results, here("regression_results", "tables", "sensitivity_analyses_table.csv"))
  write.csv(ITS_results, here("regression_results", "tables", "ITS_results_table.csv"))
}

```

# Analyses focused on private wells

```{r, echo=FALSE}

pw_results_files <- results_files[grepl("^PW", results_files)]

pw_group_results <- read.csv(here("regression_results", pw_results_files[1]))
pw_ixn_model_results <- read.csv(here("regression_results", pw_results_files[2]))

pw_group_results <- pw_group_results %>%
  dplyr::select(model_group, term, model_type, plot_estimate, nobs) %>%
  mutate(total_events = sum(dat$n_events))

pw_ixn_model_results <- pw_ixn_model_results %>%
  dplyr::select(model_group, term, model_type, plot_estimate, nobs) %>%
  mutate(total_events = sum(dat$n_events))

pw_supp_table <- rbind(pw_group_results, pw_ixn_model_results)

if (FALSE) {
  write.csv(pw_supp_table, 
            here("regression_results", "tables", "supp_table_pw.csv"))
}

```


```{r, echo=FALSE}
private_well_stats <- dat %>%
  group_by(zip3) %>%
  slice(1) %>%
  dplyr::select(zip3, weighted_percent_wells) %>%
  ungroup()

ggplot(private_well_stats) +
  geom_histogram(aes(x = weighted_percent_wells/100), bins = 20, color = "black", fill = "darkgreen") +
  scale_x_continuous(labels = scales::percent) +
  scale_y_continuous(breaks = scales::breaks_pretty()) +
  geom_vline(xintercept = 0.5, color = "darkgrey", linetype = "dashed") +
  theme_bw() +
  labs(title = "Histogram of the proportion of private well users by zip3", 
       y = "Count (n = 20)", x = "Percent residents using private wells") +
  theme(strip.background = element_rect(fill = NA),
        strip.text = element_text(face = "bold"),
        plot.title = element_text(face = "bold", hjust = 0.5),
        axis.text.x = element_text(size = 8),
        plot.subtitle = element_text( hjust = 0.5))

if (FALSE) {
  ggsave(here("figures", "private_well_histogram.png"), dpi = 400,
         width = 6)
}

```

# Hurricane Florence results

```{r, echo=FALSE}
florence_results_file <- results_files[grepl("Florence", results_files)]

florence_results <- read.csv(here("regression_results", florence_results_file))

florence_results_table <- florence_results %>%
  dplyr::select(model_group, model_type, plot_estimate, nobs)

if (FALSE) {
  write.csv(florence_results_table,
            here("regression_results", "tables", "supp_table_Florence.csv"))
}

```
